---
title: "Maps of Air Pollutants in 100 Cities "
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r message = FALSE, include = FALSE}
library(tidyverse)
library(lubridate)
library(dplyr)
library(p8105.datasets)
library(leaflet)
library(crosstalk)
library(leaflet.extras)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


The difference of daily mean AQI between 2020-02-01 ~ 2020-04-30 and 2019-02-01 ~ 2019-04-30.

```{r message = FALSE, include = FALSE}
period_19 = interval(ymd("2019-02-01"), ymd("2019-04-30"))
period_20 = interval(ymd("2020-02-01"), ymd("2020-04-30"))
```


```{r message = FALSE, include = FALSE}
all_cities_files = list.files("100_cities_data")
```


```{r message = FALSE, include = FALSE}
all_city_period_diff = 
  tibble(city = character(),
         pm25 = numeric(),
         pm10 = numeric(),
         o3 = numeric(),
         no2 = numeric(),
         so2 = numeric(),
         co = numeric(),
         )
```

```{r message = FALSE, include = FALSE}
for (city_file in all_cities_files) {
  #print(city_file)
  
  path = str_c("100_cities_data/", city_file)
  city = strsplit(city_file, split = '-')[[1]][1]
  
  cityAir = read_csv(path) %>% 
    mutate(date = as.Date(date, "%Y/%m/%d")) %>%
    arrange(date)
  
  cityAir_19 = cityAir %>% 
    filter(date %within% period_19) 
  
  cityAir_20 = cityAir %>% 
    filter(date %within% period_20)
  
  
  pm25_19 = mean(cityAir_19$pm25, na.rm = T)
  pm25_20 = mean(cityAir_20$pm25, na.rm = T)
  pm25d = pm25_20 - pm25_19
  
  pm10_19 = mean(cityAir_19$pm10, na.rm = T)
  pm10_20 = mean(cityAir_20$pm10, na.rm = T)
  pm10d = pm10_20 - pm10_19
  
  o3_19 = mean(cityAir_19$o3, na.rm = T)
  o3_20 = mean(cityAir_20$o3, na.rm = T)
  o3d = o3_20 - o3_19
  
  no2_19 = mean(cityAir_19$no2, na.rm = T)
  no2_20 = mean(cityAir_20$no2, na.rm = T)
  no2d = no2_20 - no2_19
  
  so2_19 = mean(cityAir_19$so2, na.rm = T)
  so2_20 = mean(cityAir_20$so2, na.rm = T)
  so2d = so2_20 - so2_19
  
  co_19 = mean(cityAir_19$co, na.rm = T)
  co_20 = mean(cityAir_20$co, na.rm = T)
  cod = co_20 - co_19
  
  all_city_period_diff = 
    all_city_period_diff %>% 
    add_row(city = city, 
            pm25 = pm25d,
            pm10 = pm10d,
            o3 = o3d,
            no2 = no2d,
            so2 = so2d,
            co = cod)
}

```


```{r message = FALSE, include = FALSE}
all_city_period_diff = 
  all_city_period_diff %>% 
  mutate(
    city = paste(
      toupper(substring(city, 1, 1)), 
      substring(city, 2), 
      sep = ""),
    city = fct_reorder(city, pm25, .desc = T))
```


```{r message = FALSE, include = FALSE}
cities_locations = read_csv("data/100_cities_location.csv")
all_city_period_diff = merge(all_city_period_diff, cities_locations)
```

```{r message=FALSE, include = FALSE}
long_all_city_period_diff =
  all_city_period_diff %>% 
  mutate(
    click_label = 
      str_c("<b>", paste(city, "2020 Minus 2019 Feb~Aprl Daily AQI"), 
            "</b><br>pm25 diff: ", round(pm25, digits = 2), 
            "<br>pm10 diff: ", round(pm10, digits = 2),
            "<br>o3 diff: ", round(o3, digits = 2),
            "<br>no2 diff: ", round(no2, digits = 2),
            "<br>so2 diff: ", round(so2, digits = 2),
            "<br>co diff: ", round(co, digits = 2)
            )) %>%
  pivot_longer(c(pm10, o3, no2, so2, co, pm25), names_to = "type", values_to = "diff") %>% 
  mutate(type = factor(type, levels = c("pm25", "pm10", "o3", "no2", "so2", "co")))
```


```{r message=FALSE, include=FALSE}
pal <- colorNumeric(palette = "RdBu", domain = c(-45, 45), reverse = TRUE)

allCity_sd = SharedData$new(long_all_city_period_diff)

allcity_map = allCity_sd %>% 
  leaflet(width = "100%") %>% 
  addProviderTiles(providers$Esri.NatGeoWorldMap) %>% 
  addCircleMarkers(~long, ~lat, radius = 3, opacity = 1, color = ~pal(diff), popup = ~click_label) %>% 
  addLegend("bottomleft", pal = pal, values = ~diff)
```


```{r echo=FALSE}
bscols(allcity_map)

bscols(
  filter_select("Pollutant Type", "Pollutant Type", allCity_sd, ~type, multiple = F)
)
```
Use the above drop-down selection to view different pollutants.  
  

(Note: negative values indicate improvement of air quality in 2020 compared to 2019 and positive values indicate deterioration of air quality in 2020 compared to 2019.)

The map of China included in our analysis exhibits the distribution of magnitude of air quality improvement during the lockdown period across 100 cities. Each dot represents a city we selected, and the color indicates its degree of air quality improvement. 

We can see that most cities with significant air quality improvement tend to cluster in the North China and Yangtze River Delta region. This might be partially explained by the economic structure. Industrial output value accounts for a relatively large proportion of GDP in these two areas. Normally, there are a lot of industrial activities going on there. However, during the lockdown period, nearly all of them were partially or completely suspended. As industry does have an unignorable negative impact on air quality, it’s not difficult to see why notable air quality improvement could be observed in these two regions.



