---
title: "Regression Analysis"
author: "Lin Yang"
date: "11/20/2021"
output: github_document
---

```{r, include = FALSE}
library(tidyverse)
library(ggridges)
library(modelr)
library(mgcv)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
# Regression Analysis

## Mean PM2.5 AQI Difference vs GDP and Population

### Create a data frame for regression analysis

Load and clean air quality datasets for 100 cities.
```{r, message = FALSE}
city_100_df = 
  tibble(
  file = list.files("100_cities_data")) %>% 
  mutate(
    city = str_remove(file, "-air-quality.csv"),
    path = str_c("100_cities_data/", file),
    data = map(path, read_csv)
  ) %>% 
  unnest(data) %>% 
  select(-file, -path) %>% 
  mutate(
    city = str_to_title(city),
    date = as.Date(date, format = "%Y/%m/%d"))
```

Select pm2.5 AQI during the lockdown period (Feb-Apr) and calculated mean pm2.5 AQI for both 2019 and 2020. Combine two data sets to get mean AQI differences between 2019 and 2020 for 100 cities.
```{r, message = FALSE}
pm25_2020 = 
  city_100_df %>% 
  filter(date > "2020-01-31" & date < "2020-05-01") %>% 
  group_by(city) %>%
  summarize(mean_pm25_2020 = mean(pm25, na.rm = T))
  
pm25_2019 = 
  city_100_df %>% 
  filter(date > "2019-01-31" & date < "2019-05-01") %>% 
  group_by(city) %>%
  summarize(mean_pm25_2019 = mean(pm25, na.rm = T))

pm25_diff = 
  left_join(pm25_2020, pm25_2019) %>% 
  mutate(pm25_diff = mean_pm25_2019 - mean_pm25_2020)

pm25_diff
```

Load the gdp and population dataset and join it to `pm25_diff`.

```{r, message = FALSE}
gdp_pop_df = 
  read_csv("data/gpd_and_popluation.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    gdp_trillion = gdp_billion / 1000,
    pop_million = population_thousand / 1000) %>% 
  select(city, gdp_trillion, pop_million)

diff_gdp_pop_df = 
  left_join(pm25_diff, gdp_pop_df) %>% 
  select(-mean_pm25_2020, -mean_pm25_2019)
diff_gdp_pop_df
```
We learned that air quality improvements in a city may correlate to the city's GDP and population, so we created a data frame containing mean pm2.5 AQI differences between 2019 and 2020, GDP and population in 2019 for 100 representative cities. The resulting data frame of `diff_gdp_pop_df` contains `r nrow(diff_gdp_pop_df)` observations of `r ncol(diff_gdp_pop_df)` variables. Each row represents one unique city. Below are key variables:
`city`: city name
`pm25_diff` difference of mean pm2.5 AQI during the lockdown period (Feb-Apr) between 2019 and 2020
`gdp_trillion`: 2019 GDP in trillion
`pop_million`: 2019 population in thousand


<<<<<<< HEAD
Fit linear models
=======
## Fit linear models

```{r}
fit = lm(mean_diff ~gdp_ln + pop_ln, data = diff_gdp_pop_df)
summary(fit)
=======
### Find Appropriate Transformation
>>>>>>> 31d942372fd608c8189d208d3f7241e9bac09eb9

Since the boxcox function only works with positive values for the response variable y, we removed pm25_diff less than 0 to check if a transformation is appropriate here. 
```{r, dpi = 300}
pos_diff_gdp_pop_df =
  diff_gdp_pop_df %>% 
  filter(pm25_diff > 0)
  
fit = lm(pm25_diff ~gdp_trillion + pop_million, data = pos_diff_gdp_pop_df)
MASS::boxcox(fit)
```

The box-cox plot shows that log-likelihood has the maximum value around lambda = 0.5, so square root of pm25_diff is the recommended transformation. 

### MLR
```{r}
trans_diff_gdp_pop_df =
  pos_diff_gdp_pop_df %>% 
  mutate(sqrt_pm25_diff = sqrt(pm25_diff))

trans_fit = lm(sqrt_pm25_diff ~gdp_trillion + pop_million, data = trans_diff_gdp_pop_df)

trans_fit %>% 
  broom::tidy() %>% 
  knitr::kable(caption = "Linear Regression Results")
```

After fitting a linear model for sqrt(mean pm2.5 AQI difference) dependent on gdp_trillion and pop_million, gdp_trillion variable has a slope of 0.0436 and pop_million variable has a slope of -0.0103 with p values of 0.909 and 0.846 which are extremely large. Therefore, GDP and population in a city don't have significant effects on predictions of air quality improvement, in other words, we don't have enough evidence to support that air quality improvement has a linear relationship with GDP and population. 

### Model Diagnostics


```{r, dpi = 300}
par(mfrow = c(2,2))
plot(trans_fit)
```
In residuals vs fitted plot, residuals appear to be evenly distributed around 0, indicating that residuals have constant variance. In normal QQ plot, a straight line is not seen, so our model violates the assumption that residuals are normally distributed. The scale-location plot shows that except for #11, residuals equally spread around a roughly horizontal line, confirming that residuals have constant variance. Finally, all the four plot show that there is an influential outlier labelled #11. 


### Cross Validation

Fit three models for `sqrt_pm25_diff` vs. `gdp_trillion` and `pop_million`.
```{r, warning = FALSE, dpi = 300}
nointer_linear_mod = lm(sqrt_pm25_diff ~ gdp_trillion + pop_million, data = trans_diff_gdp_pop_df)
inter_linear_mod = lm(sqrt_pm25_diff ~ gdp_trillion * pop_million, data = trans_diff_gdp_pop_df)
smooth_mod = gam(sqrt_pm25_diff ~ s(gdp_trillion, pop_million), data = trans_diff_gdp_pop_df)

trans_diff_gdp_pop_df %>% 
  gather_predictions(nointer_linear_mod, inter_linear_mod, smooth_mod) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = gdp_trillion + pop_million, y = sqrt_pm25_diff)) + 
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(~model) +
  labs(
    x = "GDP in Billion + Population in Million",
    y = "Sqrt(Mean PM2.5 AQI Difference)",
    title = "Sqrt(Mean PM2.5 AQI Difference) vs GDP and Population")
```

Cross validation for `mean_diff` vs. `gdp_trillion` and `pop_million`.
```{r, dpi = 300}
cv_df = 
  crossv_mc(trans_diff_gdp_pop_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) %>% 
  mutate(
    nointer_linear_mod  = map(train, ~lm(sqrt_pm25_diff ~ gdp_trillion + pop_million, data = .x)),
    inter_linear_mod = map(train, ~lm(sqrt_pm25_diff ~ gdp_trillion * pop_million, data = .x)),
    smooth_mod  = map(train, ~gam(sqrt_pm25_diff ~ s(gdp_trillion, pop_million), data = .x))) %>% 
  mutate(
    rmse_nointer_linear = map2_dbl(nointer_linear_mod, test, ~rmse(model = .x, data = .y)),
    rmse_inter_linear = map2_dbl(inter_linear_mod, test, ~rmse(model = .x, data = .y)),
    rmse_smooth = map2_dbl(smooth_mod, test, ~rmse(model = .x, data = .y))) 


cv_df %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_boxplot() +
  labs(
    x = "Model",
    y = "RMSE",
    title = "Distribution of RMSE across Models (Log(Mean PM2.5 AQI Difference) vs GDP +Population)") +
  theme(
    title = element_text(size = 6, face = "bold"),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10))
```

We then did cross validation for three different models of mean PM2.5 AQI difference vs. gdp_trillion and pop_million. The distribution of RMSE values for each model suggested that the smooth model worked slightly better than two linear models. There was some improvement in predictive accuracy gained by allowing non-linearity, but it was not sufficient to justify this model.

## Daily PM2.5 AQI vs Daily Average Temperature

### Create a data frame containing weather data for regression analysis

Load weather data for 10 representative cities.
```{r, message = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("CHM00054511", "CHM00058362", "CHM00050953", "CHM00054342", "CHM00055591", "CHM00056294", "CHM00056778", "CHM00059287", "CHM00057036", "CHM00057494"),
    var = c("PRCP", "TAVG"), 
    date_min = "2020-02-01",
    date_max = "2020-04-30") %>%
  mutate(
    name = recode(
      id, 
      CHM00054511 = "Beijing", 
      CHM00058362 = "Shanghai",
      CHM00050953 = "Harbin",
      CHM00054342 = "Shenyang",
      CHM00055591 = "Lhasa",
      CHM00056294 = "Chengdu",
      CHM00056778 = "Kunming",
      CHM00059287 = "Guangzhou",
      CHM00057036 = "Xian",
      CHM00057494 = "Wuhan"),
    tavg = tavg / 10,
    prcp = prcp / 10) %>%
  select(-id) %>% 
  rename(city = name) %>% 
  relocate(city)

weather_df
```

Join `weather_df` to the AQI dataset of 10 cities.
```{r, message = FALSE}
city_10_df =
  city_100_df %>% 
  filter(date > "2020-01-31" & date < "2020-05-01") %>% 
  filter(city %in% c("Beijing", "Shanghai", "Harbin", "Shenyang", "Lhasa", "Chengdu", "Kunming", "Guangzhou", "Xian", "Wuhan"))

pm25_tavg_df = 
  left_join(city_10_df, weather_df, by = c("city", "date")) %>% 
  arrange(date) %>% 
  select(city, date, pm25, tavg) %>% 
  filter(pm25 != "NA")
```

### Find appropriate transformation

```{r, dpi = 300}
fit_tavg = lm(pm25 ~tavg, data = pm25_tavg_df)
MASS::boxcox(fit_tavg)

log_pm25_tavg_df =
  pm25_tavg_df %>% 
  mutate(lnpm25 = log(pm25))

log_fit = lm(lnpm25 ~tavg, data = log_pm25_tavg_df)
log_fit %>%
  broom::tidy() %>% 
  knitr::kable(caption = "Linear Regression Results")
```

###Model diagnostics

```{r, dpi = 300}
par(mfrow = c(2,2))
plot(log_fit)
```
<<<<<<< HEAD
=======
Residuals seem to be evenly distributed around 0, which is an indication of constant variance. There are some potential outliers, such as #766 and 163. The normal QQ plot shows a roughly straight line, meaning residuals are normally distributed. Therefore, our model fitting for daily pm2.5 AQI difference dependent on daily average temperature doesn't violate assumptions on residuals. 



### Cross validation

Fit three models
```{r, dpi = 300}
linear_mod_tavg = lm(lnpm25 ~ tavg, data = log_pm25_tavg_df)
smooth_mod_tavg = gam(lnpm25 ~ s(tavg), data = log_pm25_tavg_df)
wiggly_mod_tavg = gam(lnpm25 ~ s(tavg, k = 30), sp = 10e-6, data = log_pm25_tavg_df)

log_pm25_tavg_df %>% 
  gather_predictions(linear_mod_tavg, smooth_mod_tavg, wiggly_mod_tavg) %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = tavg, y = lnpm25)) + 
  geom_point(alpha = .5) +
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(~model)
```


```{r, dpi = 300}
cv_tavg_df = 
  crossv_mc(log_pm25_tavg_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)) %>% 
  mutate(
    linear_mod  = map(train, ~lm(lnpm25 ~ tavg, data = .x)),
    smooth_mod  = map(train, ~mgcv::gam(lnpm25 ~ s(tavg), data = .x)),
    wiggly_mod  = map(train, ~gam(lnpm25 ~ s(tavg, k = 30), sp = 10e-6, data = .x))) %>% 
  mutate(
    rmse_linear = map2_dbl(linear_mod, test, ~rmse(model = .x, data = .y)),
    rmse_smooth = map2_dbl(smooth_mod, test, ~rmse(model = .x, data = .y)),
    rmse_wiggly = map2_dbl(wiggly_mod, test, ~rmse(model = .x, data = .y)))

cv_tavg_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_boxplot() +
  labs(
    x = "Model",
    y = "RMSE",
    title = "Distribution of RMSE across Models (Log(pm2.5 AQI) vs Tavg)")
```


>>>>>>> 31d942372fd608c8189d208d3f7241e9bac09eb9



