---
title: "Regression Analysis"
author: "Lin Yang"
date: "11/20/2021"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

Load and clean air quality datasets for 100 cities.
```{r, message = FALSE}
city_100_df = tibble(
  file = list.files("100_cities_data")) %>% 
  mutate(
    city = str_remove(file, "-air-quality.csv"),
    path = str_c("100_cities_data/", file),
    data = map(path, read_csv)
  ) %>% 
  unnest(data) %>% 
  select(-file, -path) %>% 
  mutate(
    city = str_to_title(city),
    date = as.Date(date, format = "%Y/%m/%d"))
```

Select pm2.5 AQI during the lockdown period (Feb-Apr) for both 2019 and 2020.
```{r}
pm25_2020 = 
  city_100_df %>% 
  filter(date > "2020-01-31" & date < "2020-05-01") %>% 
  rename(pm25_2020 = pm25) %>% 
  mutate(date = format(date, format = "%m-%d")) %>% 
  select(city, date, pm25_2020)
  
pm25_2019 = 
  city_100_df %>% 
  filter(date > "2019-01-31" & date < "2019-05-01") %>% 
  rename(pm25_2019 = pm25) %>% 
  mutate(date = format(date, format = "%m-%d")) %>% 
  select(city, date, pm25_2019)

```

Calculate daily pm2.5 AQI differences between 2019 and 2020 for each city.

```{r}
pm25_diff = 
  left_join(pm25_2020, pm25_2019, by = c("city", "date")) %>% 
  drop_na() %>% 
  mutate(pm25_diff = pm25_2019 - pm25_2020) %>% 
  group_by(city) %>% 
  summarize(mean_diff = mean(pm25_diff, na.rm = T))

pm25_diff
```


Load the gdp and population dataset and join it to `pm25_diff`.

```{r}
gdp_pop_df = 
  read_csv("data/gpd_and_popluation.csv") %>% 
  janitor::clean_names() %>% 
  mutate(pop_million = population_thousand / 1000,
         gdp_ln = log(gdp_billion, base = exp(1))) %>% 
  select(city, gdp_ln, pop_million)

df = left_join(pm25_diff, gdp_pop_df) 
```

Fit linear models
```{r}
fit = lm(mean_diff ~gdp_ln + pop_million, data = df)
summary(fit)

fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digit = 3)
```

Model Diagnostics

```{r}
df %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = gdp_ln, y = resid)) + geom_violin()

df %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = pop_million, y = resid)) + geom_violin()
```




Load weather data for 10 representative cities.
```{r, message = FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("CHM00054511", "CHM00058362", "CHM00050953", "CHM00054342", "CHM00055591", "CHM00056294", "CHM00056778", "CHM00059287", "CHM00057036", "CHM00057494"),
    var = c("PRCP", "TAVG"), 
    date_min = "2020-02-01",
    date_max = "2020-04-30") %>%
  mutate(
    name = recode(
      id, 
      CHM00054511 = "Beijing", 
      CHM00058362 = "Shanghai",
      CHM00050953 = "Harbin",
      CHM00054342 = "Shenyang",
      CHM00055591 = "Lasa",
      CHM00056294 = "Chengdu",
      CHM00056778 = "Kunming",
      CHM00059287 = "Guangzhou",
      CHM00057036 = "Xian",
      CHM00057494 = "Wuhan"),
    tavg = tavg / 10,
    prcp = prcp / 10) %>%
  select(-id) %>% 
  rename(city = name) %>% 
  relocate(city)

weather_df
```

Join `weather_df` to the AQI dataset of 10 cities and fit a linear model.
```{r}
city_10_df =
  city_100_df %>% 
  filter(date > "2020-01-31" & date < "2020-05-01") %>% 
  filter(city %in% c("Beijing", "Shanghai", "Harbin", "Shenyang", "Lasa", "Chengdu", "Kunming", "Guangzhou", "Xian", "Wuhan"))

pm25_tavg_df = 
  left_join(city_10_df, weather_df, by = c("city", "date")) %>% 
  arrange(date) %>% 
  select(city, date, pm25, tavg) %>% 
  drop_na()

fit2 = lm(pm25 ~tavg, data = pm25_tavg_df)
summary(fit2)
```



